### 2025-09-02 player-evaluation.qmd
20250902 - additions

- reaction time
- knowledge and intelligence
- work ethic and mental toughness
- incentives
    - contract performance incentives
    - whether they are in a contract year

A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
The Wonderlic test was formerly used to assess a player's cognitive abilities [@OConnell2022; archived at <https://web.archive.org/web/20220303015018/https://www.nytimes.com/2022/03/02/sports/football/nfl-wonderlic-test.html>].
A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
Contact-related incentives may lead a player to put forth greater effort.
For instance, a contract may have a performance incentive that provides a player greater compensation if they achieve a particular performance milestone (e.g., receiving yards).
Another potential incentive is if a player is in what is called their "contract year" (i.e., the last year of their current contract).
If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.

>>>>

- reaction time
- knowledge and intelligence
- work ethic and mental toughness
- competitiveness and "will to win"
- incentives
    - contract performance incentives
    - whether they are in a contract year

A player's knowledge, intelligence, and reaction time can help them gain an upper-hand even when they may not be the fastest or strongest.
The Wonderlic test was formerly used to assess a player's cognitive abilities [@OConnell2022; archived at <https://web.archive.org/web/20220303015018/https://www.nytimes.com/2022/03/02/sports/football/nfl-wonderlic-test.html>].
A player's work ethic and mental toughness may help them be resilient and persevere in the face of challenges.
The extent to which the player has a competitiveness, drive, or "will to win" may also influence their performance.
Contact-related incentives may lead a player to put forth greater effort.
For instance, a contract may have a performance incentive that provides a player greater compensation if they achieve a particular performance milestone (e.g., receiving yards).
Another potential incentive is if a player is in what is called their "contract year" (i.e., the last year of their current contract).
If a player is in the last year of their current contract, they have an incentive to perform well so they can get re-signed to a new contract.
